Neurology VI

### 2017-09-14 neuro6.rmd
sc

@@

>>>>

DDx for 1st seizure:

-  Acute vascular event/trauma: haemorrhage
-  Infn: encephalitis > meningitis
-  Systemic: ionic imbalances, hypoglycaemia, hyperthyroidism
-  Drugs, medications
-  Tumour: glioma
-  Is she pregnant?
-  Epilepsy: might not have been apparent previously

Headache from space occupying lesion: worse on lying down (sleep/morning), coughing, sneezing. Can come from tumour, abscess,..

5 most common secondaries:

1.  Breast
2.  Lung
3.  Melanoma
4.  GIT/liver

Intrinsic: glioma,...

Extrinsic: meningioma, pituitary adenoma

-  *Don't LP (yet): signs of increased ICP*
-  MRI
-  Bloods: blood glucose, FBCs (platelets, red cells), CHEM20 (ions, underlying kidney fn)
-  EEG 
-  PET later on to see if it was a secondary met

Mx:

1.  Surgical
2.  Radiotherapy
3.  Adjuvant chemotherapy: some passes BBB, some not

@@

Anatomical sites:

-  Incontinent of urine: could be generalised seizure
-  Right motor dysfn: left hemisphere

-  HPC: ever had incontinence before, trauma, how was he right before, general health in the lead up, sitting or standing, 
    -  Systems neuro: changes in vision, headaches, dizziness/faint, nausea, memory problems, concentration, personality changes
-  PMHx/PSHx: had a brain surgery before, post CVA
-  Medications/allergies/adr/drugs/cigarettes/alcohol

1.  Can't maintain airway during a seizure: oxygen
2.  Stop seizure (status epilepticus): benzodiazepine
3.  IV access

Bloods:

-  FBC: RBCs, white cells
-  ESR/CRP: inflamm
-  CHEM20: underlying kidney/liver fn, specificaly Ca and Mg
-  Drug screen
-  ABG: due to resp symptoms

-  Head CT
-  CXR
-  ECG

PDx: metastatic brain cancer

Further Ix:

-  Biopsy/chest CT to examine Ca better
-  PET

5 most common mets to brain:

-  Lung
-  Breast
-  Melanoma
-  Renal cell
-  Bowel
